Characterizing dynamical transitions by statistical complexity measures based on ordinal pattern transition networks

### 2020-11-10 manuscript_Huang.tex
updated.

\section{Choice of embedding parameters} \label{sec:embeddings}
{\color{red}We may leave out this section, if we plan a second paper which deals with embedding issue explicitly. Or we have to present the following more precisely. }
There are two important algorithmic parameters in constructing ordinal pattern transition networks, namely, embedding parameters of $D$ and $\tau$. In some cases, the choice of $D$ and $\tau$ may strongly change the results since $D$ and $\tau$ should compromise the time scales of the respective time series. In general applications of ordinal pattern analysis (i.e., permutation entropy computation $\mathcal{S}_O$), the choice of optimal $D$ and $\tau$ depends on the field of application, which requires experience of experts as systematically summarized in \cite{Riedl2013}.
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{period4_logisticExample.pdf}
\caption{\small{Example of a period-4 iterative time series ($r= 3.5$ in the Logistic map) to show the effects of embedding parameters $D$ and $\tau$. (a) iterative time series and the corresponding ordinal pattern series by different $D$ and $\tau$, (b) $D = \tau = 3$, (c) $D = \tau = 4$, (d) $D = \tau = 5$, and (e) $D = \tau = 6$. The respective permutation entropy values normalized by $\log D!$ as Eq. \eqref{eq:Ho} are $\mathcal{H}_O = 0.77$, $0$, $0.29$, and $0.11$ as indicated in the insets of each panel. (f-i) show the corresponding ordinal pattern transition networks. The weights of directed edges are suppressed due to two slightly different normalizations are introduced. Note in (g), only self-loops are available since $D = \tau = 4$ result in a unique pattern. Missing patterns are suppressed. }
\label{fig:embed}}
\end{figure*}
Here we use one example to show the important roles of $D$ and $\tau$ for a period-4 time series in Figure \ref{fig:embed}. In this example, we choose $D = \tau $ and the corresponding ordinal pattern series are significantly influenced by these parameter settings, which further result in different permutation entropy values of $\mathcal{H}_O$.  For instance, only one unique ordinal pattern is observed when $D = \tau = 4$ (Fig. \ref{fig:embed}(c, g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of one full period by the particular embedding parameters. Other values of $D$ and $\tau$ yield non-zero entropy values. The results of Fig. \ref{fig:embed} seem to be pessimistic since $D$ and $\tau$ will change the positions in the complexity-entropy planes. In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for periodic windows when the control parameter $r$ is varied. In such cases, we should focus more on the transitions between different dynamic regimes. Instead, we show that the results are robust when the control parameters are systematically changed while using the same parameter settings of $D$ and $\tau$.

>>>>

\section{Choice of embedding parameters} \label{sec:embeddings}
There are two important algorithmic parameters in constructing ordinal pattern transition networks, namely, embedding parameters of $D$ and $\tau$. In some cases, the choice of $D$ and $\tau$ may strongly change the results since $D$ and $\tau$ should compromise the time scales of the respective time series. In general applications of ordinal pattern analysis (i.e., permutation entropy computation $\mathcal{S}_O$), the choice of optimal $D$ and $\tau$ depends on the field of application, which requires experience of experts as systematically summarized in \cite{Riedl2013}.
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{period4_logisticExample.pdf}
\caption{\small{Example of a period-4 iterative time series ($r= 3.5$ in the Logistic map) to show the effects of embedding parameters $D$ and $\tau$. (a) iterative time series and the corresponding ordinal pattern series by different $D$ and $\tau$, (b) $D = \tau = 3$, (c) $D = \tau = 4$, (d) $D = \tau = 5$, and (e) $D = \tau = 6$. The respective permutation entropy values normalized by $\log D!$ as Eq. \eqref{eq:Ho} are $\mathcal{H}_O = 0.77$, $0$, $0.29$, and $0.11$ as indicated in the insets of each panel. (f-i) show the corresponding ordinal pattern transition networks. The weights of directed edges are suppressed due to two slightly different normalizations are introduced. Note in (g), only self-loops are available since $D = \tau = 4$ result in a unique pattern. Missing patterns are suppressed. }
\label{fig:embed}}
\end{figure*}
Here we use one example to show the important roles of $D$ and $\tau$ for a period-4 time series in Figure \ref{fig:embed}. In this example, we choose $D = \tau $ and the corresponding ordinal pattern series are significantly influenced by these parameter settings, which further result in different permutation entropy values of $\mathcal{H}_O$.  For instance, only one unique ordinal pattern is observed when $D = \tau = 4$ (Fig. \ref{fig:embed}(c, g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of one full period by the particular embedding parameters. Other values of $D$ and $\tau$ yield non-zero entropy values. The results of Fig. \ref{fig:embed} seem to be pessimistic since $D$ and $\tau$ will change the positions in the complexity-entropy planes. In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for periodic windows when the control parameter $r$ is varied. In such cases, we should focus more on the transitions between different dynamic regimes. Instead, we show that the results are robust when the control parameters are systematically changed while using the same parameter settings of $D$ and $\tau$.
